Inference of Gravitational Lens Time Delays Using Gaussian Processes

### 2012-05-30 doc/ms.tex
Removed a bunch of old stuff (it is still accessible from the history, don't worry!)

\input{macros.tex}

@@

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{A Sample of Mock Lenses}

Motivated by the need to accurately measure time delays for lens cosmography
in the multi-filter optical imaging survey era, we  we selected three
quadruply-imaged quasar lens systems from the mock LSST  catalog provided by
\citet[][hereafter OM10]{OM10}.\footnote{The catalog is available from
\texttt{http://kipac.stanford.edu/collab/research/lensing}}

The OM10 mock lenses are designed to represent massive elliptical galaxies;
they have singular isothermal mass distributions, with velocity dispersions
$\vdisp$ drawn from the SDSS velocity function \citep[extrapolated to higher
redshifts assuming no evolution, as in][]{Ogu++08}. The sources were drawn
from a model quasar luminosity function that matches the observed bright and
faint ends. External shears, drawn from a plausible distribution, were
applied. The mock catalog is the product of performing the lens abundance
calculation for the LSST survey by Monte Carlo integration; OM10 found some
3000 well-measured lens systems to be detectable at LSST single-epoch depth
and image quality. Here we consider a typical slice through this sample: from
the 22 systems lying between $0.4 < \zd < 0.6$, $1.6 < \zs < 2.4$ and $150 <
\vdisp < 300$\kms, three were chosen as being representative of the possible
image configurations while also being particularly prone to microlensing
effects: 6 out of the 12 images lie within the lens galaxy effective radius.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{minipage}{0.32\linewidth}
\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=3042348.eps}
\end{minipage}
\begin{minipage}{0.32\linewidth}
\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=3424637.eps}
\end{minipage}
\begin{minipage}{0.32\linewidth}
\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=5058554.eps}
\end{minipage}\hfill
\caption{The three simulated lenses, showing image positions (blue) relative
to the lens light effective area (orange). The fraction of surface mass
density in stars ($f^{*}$) at each image position is shown under each image's\label{lenslayouts}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
\begin{center}
\begin{tabular}{ccc}
\hline
Parameter      & Possible values \\
\hline\hline
$tau_q$/days   & 31.62, {\bf 316.2}, 3162  \\ 
$sigma_q$      & 0.004, {\bf 0.008}, 0.016 \\ 
\hline
\end{tabular}
\end{center}
\caption{Simulated source parameters. The damped random walk variability model
is controlled by two parameters, $tau_q$ and $sigma_q$; the values used here
are representative of those observed in the SDSS stripe 82 spectroscopic 
quasar sample. Our fiducial source lies at the centre of the resulting 3 by 3
grid, and its variability parameters are highlighted in bold font.
\label{tab:simsrc}}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We then apply the appropriate magnifications and time delays to generate
4-year lightcurves for each image of our three mock lenses. We sample the
lightcurves in several different ways, to explore the robustness of our time
delay inference. Keeping the monitoring campaign length fixed at 4 years, we
down-sample to a plausible cadence (separation between successive
observations) and season length. Again, we choose representative values:
COSMOGRAIL \citep{cosmograil} typically monitor lenses for 8 months per year
at a cadence of 4 days, LSST will likely have seasons closer to 4 months, with
a single-filter cadence closer to 8 days (although this is somewhat
pessimistic). 4-month seasons with 4 day cadence mimics LSST monitoring should
multi-filter joint analysis be possible, while 8-month seasons with 8-day
cadence represents a cheaper version of a dedicated monitoring campaign. These
choices resulted in 4 possible observing strategies, with total numbers of
measurements of 60, 120, 120 and 240. A more thorough investigation of time
delay measurement feasibility with LSST is left to future work. Four observing
strategies, times 3 times 3 possible variable sources, lensed by 3 deflectors 
makes a total of 108 simulated datasets, and 324 independent time delays to be
inferred.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table*}
\begin{center}
\begin{tabular}{ccccc}
\hline
Observing strategy & Cadence (days) & Season length (days) 
& Epochs/season & Total epochs\\
\hline\hline
A  & 8  & 120 & 15  &  60 \\ 
\textbf{B}  & \textbf{4}  & \textbf{120} & \textbf{30}  & \textbf{120} \\
C  & 8  & 240 & 30  & 120 \\
D  & 4  & 240 & 60  & 240 \\
\hline
\end{tabular}
\end{center}
\caption{Simulating observing strategies explored. 4 day cadence and 8 month
seasons are an approximation to those used in the COSMOGRAIL survey, while
LSST will provide closer to 8 day cadence and 4 month seasons. Strategy B is
our fiducial one, corresponding approximately to (somehow) using multiple
LSST filters to achieve a 4-day cadence over four consecutive 4-month seasons.
\label{observing}}
\end{table*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\TODO{PJM}{Check these  numbers, esp COSMOGRAIL...}

Two example simulated datasets are shown in Figure~\ref{fig:simdata}. In  the
next sub-section we present results from our Gaussian Process modelling of all 108
datasets.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{minipage}{0.48\linewidth}
\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=3042348_tau=31_sigma=4_cadence=4_season=8_lightcurve.eps}
\end{minipage}
\begin{minipage}{0.48\linewidth}
\centering\includegraphics[width=\linewidth]{figs1/oguri_qso_ID=5058554_tau=3162_sigma=16_cadence=8_season=4_lightcurve.eps}
\end{minipage}\hfill
\caption{Two example simulated lensed AGN datasets (with no microlensing).
Left: a minimally-variable source ($\tau_q = 31.62$ days, $\sigma_q = 0.004$),
lensed into a cross configuration (lens ID 3042348), and monitored with high
cadence and long season length (4 days, 8 months, strategy D). Right: a
maaximally-variable source ($\tau_q = 3162$ days, $\sigma_q = 0.016$), lensed
into a major axis cusp configuration (lens ID 5058554), and monitored with low
cadence and short season length (8 days, 4 months, strategy A).
\label{fig:simdata}}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Results}

Strategy B, dt recovery across sample. 3x3x3 = 27 inferences, 27x18 dt estimates. Are
uncertainties on dt accurate? Trends with dt.

Effect of changing observing strategy.

Mean mag and var par recovery across sample. Trends with tau and sigma.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

@@

\TODO{GD}{Write this section please!}
\TODO{PJM/BJB}{Suitable illustrative figure?}

Stellar mass distributions were ``painted on'' to the model lens galaxies, as
described in Dobler et al.\, (in preparation). Established galaxy scaling
relations were used to draw luminosities and effective radii given the model
velocity dispersions; a surface brightness distribution with the same
ellipticity as that of the lens mass, and with a de Vaucoleurs cite profile 
were assumed. Finally, a mass to light ratio was assumed, and the fraction of
surface mass density in the form of stars ($\fstar$) 
computed, at each image position. This parameter, the local lens
convergence, and its shear, were then used to generate a starfield at each
image position, and compute the corresponding source plane magnification map,
using the code of ...

Wambsganss maps. Then, parameters chosen to span the possibilities. Constant
speed traversals, 2 speeds (fast and slow).  2 source sizes, from Kochanek etc
work, separated by factor of 3 (large and small).  fstar values chosen in
catalog and in lens selection, check they are high (c. 0.5).  Check timescales
compared to length of dataset. Illustrative plot.

Apply to 3 lenses with "typical" intrinsic variabilities, ie the middle tau
and sigma of the 3x3 grid in the previous section.

\subsubsection{Results}

dt recovery across sample. 3x(2x2) = 12 inferences, 12x18=216 dt estimates.
Are uncertainties on dt accurate? Trends with dt.

Mean mag and ml par recovery across sample. 48 microlensing lightcurves (4
images times 3 lenses x 4 setups). Trends with fstar, source size and v.

% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -

\subsection{Discussion}

>>>>

%\input{macros.tex}

@@

@@

@@

\subsubsection{Methodology}
\subsection{Discussion}
